---
title: "Homework #7 2.1-2.6: Nontransitive Dice"
author: "Andrei Dominic Magno"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Simulate in Rstudio the nontransitive dice probabilities in Exercise 2.6. Upload a file with your code and output.

Reference: (Consider three nonstandard dice. Instead of the numbers
1 through 6, die A has two 3’s, two 5’s, and two 7’s; die B has two
2’s, two 4’s, and two 9’s; and die C has two 1’s, two 6’s, and two 8’s, as in
Figure 2.8.

Suppose dice A and B are rolled. Show that A is more likely to get the higher
number. That is, P(A > B) > 0.50, where {A > B} denotes the event that
A beats B. Hint: Condition on the outcome of die A.
Now show that if B and C are rolled, B is more likely to get the higher number.
And, remarkably, if C and A are rolled, C is more likely to get the higher
number.

```{r Nontransitive Dice; A beats B}
#A beats B
die_A <- c(3, 3, 5, 5, 7, 7)
die_B <- c(2, 2, 4, 4, 9, 9)

sim_A_greater_B = function()
{
num_A = sample(die_A, 1, replace = TRUE)
num_B = sample(die_B, 1, replace = TRUE)

if (num_A > num_B) 1 else 0
}

simlist = replicate(10000, sim_A_greater_B())
mean(simlist)
```
```{r Nontransitive Dice; B beats C}
#B beats C
die_B <- c(2, 2, 4, 4, 9, 9)
die_C <- c(1, 1, 6, 6, 8, 8)

sim_B_greater_C = function()
{
num_B = sample(die_B, 1, replace = TRUE)
num_C = sample(die_C, 1, replace = TRUE)

if (num_B > num_C) 1 else 0
}

simlist = replicate(10000, sim_B_greater_C())
mean(simlist)
```

```{r Nontransitive Dice; C beats A}
#C beats A
die_A <- c(3, 3, 5, 5, 7, 7)
die_C <- c(1, 1, 6, 6, 8, 8)

sim_C_greater_A = function()
{
num_C = sample(die_C, 1, replace = TRUE)
num_A = sample(die_A, 1, replace = TRUE)

if (num_C > num_A) 1 else 0
}

simlist = replicate(10000, sim_C_greater_A())
mean(simlist)
```


Many relationships in life are transitive. For instance, if Amy is taller than
Ben and Ben is taller than Charlie, then Amy is taller than Charlie. But
these dice show that the relation “more likely to roll a higher number” is not
transitive.

The dice are the basis of a magic trick. You pick any die. Then I can always
pick a die that is more likely to beat yours. If you pick A, I pick C. If you pick
B, I pick A. And if you pick C, I pick B.)


