---
title: 'Homework #6 1.9'
author: "Andrei Dominic Magno"
date: "2026-02-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Homework 4
1.41 Modify the code in the R script CoinFlip.R to simulate the probability of getting exactly
one head in four coin tosses.

```{r one head in four coin tosses}
#coinFlip.R

n <- 10000
simlist <- numeric(n)
for (i in 1:n) {
  trial <- sample(0:1, 4, replace = TRUE)  # 4 coin tosses
  success <- if (sum(trial) == 1) 1 else 0 # exactly one head
  simlist[i] <- success
}

mean(simlist)
```


1.43 Use R to simulate the probability of getting at least 8 in the sum of two dice rolls.
```{r at least 8 in the sum of two dice rolls}

n <- 10000
simlist <- numeric(n)
for (i in 1:n) {
  trial <- sample(1:6, 2, replace = TRUE)  # two dice rolls
  success <- if (sum(trial) >= 8) 1 else 0 #at least 8
  simlist[i] <- success
}

mean(simlist)
```

1.57
Judith has a penny, nickel, dime, and quarter in her pocket. So does Kory. They both reach into
their pockets and choose a coin. Let upper X be the greater (in cents) of the two.
Assume that coin selections are equally likely.
Simulate to find the probability that Judith's coin is worth more than Kory's? (It is not 1 slash 2
period)
```{r probability judith > kory}

n <- 10000
simlist <- numeric(n)

coins <- c(1,5,10,25) 
for (i in 1:n) {
  judith <- sample(coins, 1)  #judith's pocket
  kory <- sample(coins, 1) #kory's pocket
  success <- if (judith > kory) 1 else 0 
  simlist[i] <- success
}

mean(simlist)
```
