---
title: "R Coding Assignment 5"
author: "Andrei Dominic Magno"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R Coding Assignment 5

1. Birth weights of babies in the United States can be modeled by a normal distribution with mean 3250 grams and standard deviation 550 grams. 
Those weighing less than 2500 grams are considered to be of low birth weight.

A. What proportion of babies would the normal distribution predict as weighing
more than 10 pounds (4536 grams) at birth? You should always start with a
sketch of the area that you are looking for!
B. Determine the probability that a randomly selected baby weighs between 3000
and 4000 grams at birth.
C. How little would a baby have to weigh to be among the lightest 2.5% of all
newborns?
D. How much would a baby have to weigh to be among the heaviest 10% of all
newborns?

``` {r Q1}
cat("A) The proportion of babies that the normal distribution would predict as weighing more than 10 pounds (4536 grams) at birth is", round(1 - pnorm(4536, 3250, 550), 4), '\n')
cat("B) The probability that a randomly selected baby weighs between 3000 and 4000 grams at birth is", round(pnorm(4000, 3250, 550) - (pnorm(3000, 3250, 550)), 4), '\n')
cat("C) A baby have to weigh",qnorm(0.025, 3250, 550),"grams to be among the lightest 2.5% of all newborns", '\n')
cat("D) a baby have to weigh", round(qnorm(0.9, 3250, 550), 2), "grams to be among the heaviest 10% of all newborns", '\n')
```

2. The weekly salaries of teachers in one state are normally distributed with a mean of
$490 and a standard deviation of $45. What is the probability that a randomly
selected teacher earns more than $525 a week?

``` {r Q2}
cat("The probability that a randomly selected teacher earns more than $525 a week is", round(1 - pnorm(525, 490, 45), 4), '\n')
```

3. The number of days ahead travelers purchase their airline tickets can be modeled
by an exponential distribution with the average amount of time equal to 15 days.
Find the probability that a traveler will purchase a ticket fewer than ten days in
advance.

``` {r Q3}
lambda = 1 / 15
cat("The probability that a traveler will purchase a ticket fewer than ten days in advance is", round(pexp(10, lambda), 4), '\n')
```


4. Packages have a nominal net weight of 1000 g. However their actual net weights have a
uniform distribution over the interval 980 g to 1030 g.
(a) Find the probability that the net weight of a package is less than 1000 g.
(b) Find the probability that the net weight of a package is less than w g, where
990 < w < 1030.

``` {r Q4}
cat("A) The probability that the net weight of a package is less than 1000 g is", round(punif(1000, 980, 1030), 2), '\n')
cat("B) The probability that the net weight of a package is less than w g, where 990 < w < 1030 is, 'P(X < w) = (w - 980) / 50' \n")
cat("   For example, if w = 1000, the probability is", (1000 - 980) / 50, '\n')
```