---
title: "R code Assingment 1"
author: "Andrei Dominic Magno"
date: "2026-01-30"
output:
  word_document: default
  pdf_document: default
---


### R A First Look at Simulation

This section of the text introduces the concept of Monte Carlo simulation, along with some R commands using two examples (which are provided below or as script files). The first code starts with Example 1.32.

```{r example 1}
n <- 100 #set for example
sample(0:1, n, replace = TRUE)
mean(sample(0:1, n, replace = TRUE))
```

### try creating a sample now

use the code above and make a sample of sampling from the digits 0-9 and find the mean.

```{r Code here 1}

```

###Here are some of the functions that will be useful as we go through the course.

choose(n,k) - computes combination of choosing k objects from n.
prod(n:m) - computes the product of values from n to m. To do a permutation of r objects out of n,
you want to use prod(n:n-r+1). (For whatever reason, this doesn’t work if you assign n and r first, just
put in the numbers).
factorial(x) - computes x!


```{r example 2}
choose(n = 10, k = 2)
## [1] 45
prod(8:6) # permutation of 3 objects out of 8
## [1] 336
factorial(x = 4)
##[1] 24
```
## Try the following
Find how many ways you can select 10 things from a group of 30

Find the total possible ways to arrange 10 people in a line for a photo.


Find 9!
```{r practice}

```

