---
title: 'R Coding Assignment #3'
author: "Andrei Dominic Magno"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. If 10% of people who are given a certain drug experience Dizziness, find these
probabilities using R code for a sample of 25 people who take the drug.
A. Exactly 3 people will become dizzy
B. At least 9 people will become dizzy.
C. Less than 7 people will become dizzy.

``` {r Q1a}
p <- 0.1
n <- 25

cat("The probability that exactly 3 people will become dizzy is",dbinom(3, n, p), "\n")
cat("The probability that at least 9 people will become dizzy is", 1 - pbinom(8, n, p), "\n")
cat("The probability that less than 7 people will become dizzy is", pbinom(6, n, p), "\n")

```

2. There were ten green bottles sitting on the wall. The probability of a green bottle
accidentally falling is 0.95. What is the probability that fewer than 8 of the green
bottles accidentally fall?

``` {r Q2}
p <- 0.95
n <- 10
cat("The probability that fewer than 8 of the green bottles accidentally fall is", pbinom(7, n, p), "\n")
```

3. The LMB Company manufactures tires. They claim that only .007 of LMB tires
are defective. What is the probability of finding 2 defective tires in a random
sample of 50 LMB tires?

``` {r Q3}
p <- 0.007
n <- 50
cat("The probability of finding 2 defective tires in a random sample of 50 LMB tires is",dbinom(2, n, p), "\n")
```