---
title: "R coding Assignment 4 - Andrei Dominic Magno"
author: "Andrei Dominic Magno"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R coding Assignment 4

1. Traffic accidents at a particular intersection follow Poisson distribution with an
average rate of 1.4 per week.
a) What is the probability that the next week is accident-free?
b) What is the probability that there will be exactly 3 accidents next week?
c) What is the probability that there will be at most 2 accidents next week?
d) What is the probability that there will be at least 2 accidents during the next
two weeks?
e) What is the probability that there will be exactly 5 accidents during the next
four weeks?
f) What is the probability that there will be exactly 2 accidents tomorrow?
g) What is the probability that the next accident will not occur for three days?

``` {r Q1}
m <- 1.4
cat("a) The probability that the next week is accident-free is", dpois(0, m), "\n")
cat("b) The probability that there will be exactly 3 accidents next week is", dpois(3, m), "\n")
cat("c) The probability that there will be at most 2 accidents next week is", ppois(2, m), "\n")
cat("d) The probability that there will be at least 2 accidents during the next two weeks is", 1 - ppois(2 - 1, m * 2), "\n")
cat("e) The probability that there will be exactly 5 accidents during the next four weeks is", dpois(5, m * 4), "\n")
cat("f) The probability that there will be exactly 2 accidents tomorrow is", dpois(2, m/7), "\n")
cat("g) The probability that the next accident will not occur for three days is", dpois(0, (m/7 * 3)), "\n")

```

2. In Major League Baseball’s Home Run Derby, each contestant is allowed to
keep swinging the bat until they have made 10 “outs”. (An “out” is anything that
is not a home run.) If Barry Bonds has a 70% chance of hitting a home run on
any given swing, what is the probability that he hits at least 10 home runs
before his turn is up?

``` {r Q2}
p <- 0.7
outs <- 10
cat("The probability that he hits at least 10 home runs before his turn is up is", 1 - pnbinom(9, 10, 1-p), "\n")
```

3. A quarterback completes 44% of his passes. We want to observe this
quarterback during one game to see how many pass attempts he makes before
completing one pass.
A. What is the probability that the quarterback throws 3 incomplete passes before
he has a completion?
B. Determine the probability that it takes more than 5 attempts before he completes
a pass.

``` {r Q3}
p <- 0.44
cat("A) The probability that the quarterback throws 3 incomplete passes before he has a completion is", dgeom(3, p), "\n")
cat("B) The probability that it takes more than 5 attempts before he completes a pass is", 1 - pgeom(4, p), "\n")
```

4. Suppose a telemarketer has a 15% chance of making a sale on any given
phone call. He is required to make 10 successful sales before leaving for the
day. What is the probability that he needs to make more than 40 calls?

``` {r Q4}
p <- 0.15
cat("The probability that he needs to make more than 40 calls is", 1- pnbinom(40 - 10, 10, p))
```

5. A company (the producer) supplies microprocessors to a manufacturer (the
consumer) of electronic equipment. The microprocessors are supplied in batches
of 50. The consumer regards a batch as acceptable provided that there are not
more than 5 defective microprocessors in the batch. Rather than test all of the
microprocessors in the batch, 10 are selected at random and tested.
A. Find the probability that out of a sample of 10, d = 0, 1, 2, 3, 4, 5 are
defective when there are actually 5 defective microprocessors in the batch.
B. Suppose that the consumer will accept the batch provided that not more than
m=2 defectives are found in the sample of 10.
  i. Find the probability that the batch is accepted when there are 5
defectives in the batch.
  ii. Find the probability that the batch is rejected when there are 3
defectives in the batch.

``` {r Q5}
N <- 50
n <- 10
cat("A) The probability that out of a sample of 10, d = 0, 1, 2, 3, 4, 5 are defective when there are actually 5 defective microprocessors in the batch is: \n")
cat("P(X=0):", dhyper(0, 5, 45, 10), "\n")
cat("P(X=1):", dhyper(1, 5, 45, 10), "\n")
cat("P(X=2):", dhyper(2, 5, 45, 10), "\n")
cat("P(X=3):", dhyper(3, 5, 45, 10), "\n")
cat("P(X=4):", dhyper(4, 5, 45, 10), "\n")
cat("P(X=5):", dhyper(5, 5, 45, 10), "\n")
cat("Bi) the probability that the batch is accepted when there are 5 defectives in the batch", phyper(2, 5, 45, 10), "\n")
cat("Bii) the probability that the batch is accepted when there are 3 defectives in the batch", 1 - phyper(2, 3, 47, 10), "\n")
```


6. In a certain town, 40% of the eligible voters prefer candidate A, 10% prefer
candidate B, and the remaining 50% have no preference. You randomly sample
10 eligible voters. What is the probability that 4 will prefer candidate A, 1 will
prefer candidate B, and the remaining 5 will have no preference?

``` {r Q6}
cat("The probability that 4 will prefer candidate A, 1 will prefer candidate B, and the remaining 5 will have no preference is", dmultinom(c(4,1,5), prob = c(0.4,0.1,0.5)))
```
